```{r, echo = FALSE}
NCORES = 1L
CACHE_DIR = "pomp_cache/"
run_level = 2
NP = switch(run_level, 50, 1e3, 3e3) # number of particles
NMIF_S = switch(run_level, 5, 50, 100) # number of filtering iterations - small
NMIF_L = switch(run_level, 10, 100, 200) # - large
NREPS_EVAL = switch(run_level, 5, 20, 40) # number of replications in likelihood evaluation
NREPS_LOCAL = switch(run_level, 10, 20, 30) # number of replications in local search
NSTART = switch(run_level, 50, 500, 800) # number of starting points in the global search
NSIM = switch(run_level, 50, 100, 500) # number of simulations

PARAMS_FILE = sprintf("seir_params_runlevel=%i.csv", run_level)

suppressPackageStartupMessages({
  library(foreach)
  library(doParallel)
  library(doRNG)
  library(tidyverse)
  library(pomp)
})
cl = makeCluster(NCORES)
registerDoParallel(cl)
registerDoRNG(625904618)
```

## SEIR Model

### Model specification

```{r}
cases = read.csv("cases_deaths_by_county_date.csv",
                 colClasses = list(Date = "Date")) %>%
  # select records of confirmed cases in Washtenaw County
  filter(!is.na(Date), CASE_STATUS == "Confirmed", COUNTY == "Washtenaw") %>%
  select(-Updated, -CASE_STATUS, -COUNTY, -ends_with("Cumulative")) %>%
  # select records in 2020 only
  filter(Date < as.Date("2021-01-01")) %>%
  # transform `Date` to numeric form
  mutate(Time = 1:n())
covid_t0 = min(cases$Time)
```

[TODO: Add SEIR model settings, justifications for the beta step function, and justification for the initial state values (E and I).]{.comment}

```{r}
seir_step = Csnippet("
  double Beta;
  if(intervention == 1) Beta = b1;
  else if(intervention == 2) Beta = b2;
  else if(intervention == 3) Beta = b3;
  else if(intervention == 4) Beta = b4;
  else Beta = b5;
  
  double dN_SE = rbinom(S, 1 - exp(-Beta * I / N * dt));
  double dN_EI = rbinom(E, 1 - exp(-mu_EI * dt));
  double dN_IR = rbinom(I, 1 - exp(-mu_IR * dt));
  S -= dN_SE;
  E += dN_SE - dN_EI;
  I += dN_EI - dN_IR;
  H += dN_IR;
")

seir_rinit = Csnippet("
  S = nearbyint(eta * N);
  E = 100;
  I = 200;
  H = 0;
")

dmeas <- Csnippet("
  double tol=1.0e-25;
  double mean =rho*H;
  double sd =sqrt(pow(tau*H,2)+rho*H);
  if(Cases>0.0){
    lik=pnorm(Cases+0.5,mean,sd,1,0)-pnorm(Cases-0.5,mean,sd,1,0)+tol;
  } else {
    lik=pnorm(Cases+0.5,mean,sd,1,0)+tol;
  }
  if(give_log) lik=log(lik);
")

rmeas <- Csnippet("
  Cases = rnorm(rho*H, sqrt(pow(tau*H,2)+rho*H));
  if(Cases>0.0){
    Cases=nearbyint(Cases);
  } else {
    Cases=0.0;
  }
")

seir_covar <- covariate_table(
  t = cases$Time,
  intervention = c(rep(1, 23),
                   rep(2, 77),
                   rep(3, 20),
                   rep(4, 75),
                   rep(5, 111)),
  times = "t")

covidSEIR = cases %>% select(Time, Cases) %>%
  pomp(
    times = "Time", t0 = covid_t0,
    rprocess = euler(seir_step, delta.t = 1), # delta.t set to 1 day
    rinit = seir_rinit,
    rmeasure = rmeas,
    dmeasure = dmeas,
    accumvars = "H",
    partrans=parameter_trans(
      log = c("mu_EI", "mu_IR", "tau", "b1", "b2", "b3", "b4", "b5"),
      logit = c("rho", "eta")
    ),
    statenames = c("S", "E", "I", "H"),
    paramnames = c("b1", "b2", "b3", "b4", "b5", "mu_EI", "mu_IR", 
                   "eta", "rho", "N", "tau"),
    covar = seir_covar
  )
```

### Choosing starting points

As of the [2019 census](https://fred.stlouisfed.org/series/MIWASH1POP), the resident population in Washtenaw County was 367,601. We will use this value as an approximation for the population in March 2020. In the following analysis, we will fix the population $N$ to 367,601.  

According to the Centers for Disease Control and Prevention (CDC), the incubation period (the time from exposure to development of symptoms) of coronaviruses ranges from 2–14 days. [[2]](https://www.cdc.gov/coronavirus/2019-ncov/hcp/faq.html#Transmission)
So we expect $\mu_{EI}$ to fall into the range of $\frac{1}{2\text{ wk} = 14\text{ da}} = 0.071\text{ da}^{-1}$ to $\frac{1}{2\text{ day}} = 0.5\text{ da}^{-1}$.
It can take at least two weeks for people to recover from COVID-19 infection. [[3]](https://www.reuters.com/article/us-health-coronavirus-recovery-idUSKCN24P2DC). So we can expect $\mu_{IR}$ to fall into the range of $0.048-0.071\text{ da}^{-1}$. For the starting values, we will set both $mu_EI$ and $mu_IR$ to 0.1.

[TODO: Add some more analysis about eta and rho (and tau?)]{.comment}

Based on the above information, we can specify the starting points of the local search: 

\[
\begin{cases}
b_1 = 3,\ b_2 = 0.5,\ b_3= 4,\ b_4 = 1.5,\ b_5 = 3.2 \\
\mu_{EI} = 0.1 \\
\mu_{IR} = 0.1 \\
\rho = 0.5 \\
\eta = 0.09 \\
\tau = 0.001 \\
N = 367,601
\end{cases}
\]

```{r}
pop_washtenaw = 367601
params = c(b1 = 3, b2 = 0.5, b3 = 4, b4 = 1.5, b5 = 3.2, 
            mu_EI = 0.1, mu_IR = 0.1, rho = 0.5, eta = 0.09, 
            tau = 0.001, N = pop_washtenaw)
fixed_params = params[c("N")]
params_rw.sd = rw.sd(b1 = 0.02, b2 = 0.02, b3 = 0.02, b4 = 0.02, b5 = 0.02, 
                     mu_EI = 0.02, mu_IR = 0.02, rho = 0.02, 
                     tau = 0.0001, eta = ivp(0.02))
```


```{r}
registerDoRNG(1235252)
bake(file = sprintf("%srunlevel=%i_%s", CACHE_DIR, run_level, "lik_starting_values.rds"), {
  foreach(i=1:10, .combine = c) %dopar% {
    library(pomp)
    covidSEIR %>% pfilter(params=params,  Np=500)
  }
}) -> pf
pf %>% logLik() %>% logmeanexp(se = TRUE)
```

The simulation based on the starting values has an outbreak but does not capture the data well. Next, we will use iterative filtering to search for the maximum likelihood estimates (MLE). 

```{r, echo = FALSE, fig.width = 6, fig.height = 3}
plot_simulation = function(sim_dat) {
  sim_dat %>%
    ggplot() +
    theme_bw() +
    geom_line(aes(Time, Cases, group = .id, 
                  color = (.id == "data"), alpha = (.id == "data"), 
                  linetype = (.id == "data"))) +
    scale_color_manual(values = c("#18bc9c", "#c61919")) +
    scale_alpha_manual(values = c(0.5, 1)) +
    scale_linetype_manual(values = c(5, 1)) +
    guides(color = FALSE, linetype = FALSE, alpha = FALSE)
} # plot_simulation()

covidSEIR %>%
  simulate(params = params, nsim = 20, format = "data.frame", include.data = TRUE) %>% 
  plot_simulation()
```

### Local Search

```{r, fig.height = 5, fig.width = 8}
run_id = 1
registerDoRNG(482947940)
bake(file = sprintf("%srunlevel=%i_%s", CACHE_DIR, run_level, "local_search.rds"),{
  foreach(i = 1:NREPS_LOCAL, .combine = c) %dopar% {
    suppressPackageStartupMessages({
      library(tidyverse)
      library(pomp)
    })
    covidSEIR %>%
      mif2(
        params = params,
        Np = NP, Nmif = NMIF_S,
        cooling.fraction.50 = 0.5,
        rw.sd = params_rw.sd
      )
  } -> mifs_local
  attr(mifs_local,"ncpu") <- getDoParWorkers()
  mifs_local
}) -> mifs_local
t_loc <- attr(mifs_local,"system.time")
ncpu_loc <- attr(mifs_local,"ncpu")
cat(sprintf("Local search (iterative filtering) finished in %4.3f seconds\n", t_loc["elapsed"]))

mifs_local %>%
  traces() %>%
  melt() %>%
  ggplot(aes(x = iteration, y = value, group = L1, color = factor(L1))) +
  theme_bw() +
  geom_line() +
  guides(color = FALSE) +
  facet_wrap(~variable, scales = "free_y")
```

```{r}
# likelihood est. for local search results
registerDoRNG(900242057)
bake(file = sprintf("%srunlevel=%i_%s", CACHE_DIR, run_level, "lik_local.rds"),{
  foreach(mf = mifs_local, .combine = rbind) %dopar% {
    suppressPackageStartupMessages({
      library(tidyverse)
      library(pomp)
    })
    ll = replicate(NREPS_EVAL, logLik(pfilter(mf, Np = NP))) %>% 
         logmeanexp(se = TRUE)
    coef(mf) %>% bind_rows() %>% bind_cols(loglik = ll[1], loglik.se = ll[2])
  } -> results
  attr(results,"ncpu") <- getDoParWorkers()
  results
}) -> results
t_local <- attr(results,"system.time")
ncpu_local <- attr(results,"ncpu")
cat(sprintf("Local search (likelihood estimation) finished in %4.3f seconds\n", t_local["elapsed"]))
```

```{r}
results %>% arrange(-loglik) %>% head %>% knitr::kable(digits = 3)
```

```{r, fig.height = 7, fig.width = 7}
pairs(~loglik + b1 + b2 + b3 + b4 + b5, data = results, pch = 16)
pairs(~loglik + mu_EI + mu_IR + eta + rho + tau, data = results, pch = 16)
```

### Global Search

```{r}
run_id = 2

# create a box of starting values (for parameters)
set.seed(2062379496)
guesses = runif_design(
  lower = c(b1 = 0, b2 = 0, b3 = 0, b4 = 0, b5 = 0, 
            mu_EI = 0, mu_IR = 0, rho = 0, eta = 0, tau = 0),
  upper = c(b1 = 10, b2 = 10, b3 = 10, b4 = 10, b5 = 10, 
            mu_EI = 1, mu_IR = 1, rho = 1, eta = 0.3, tau = 0.1),
  nseq = NSTART
)

mf1 = mifs_local[[1]] # take the output of previous IF process (local search)

bake(file = sprintf("%srunlevel=%i_%s", CACHE_DIR, run_level, "global_search.rds"),{
  registerDoRNG(1270401374)
  foreach(guess=iter(guesses, "row"), .combine = rbind) %dopar% {
    suppressPackageStartupMessages({
      library(tidyverse)
      library(pomp)
    })
    mf = mf1 %>% # cooling.fraction.50 = 0.5
      mif2(params = c(unlist(guess), fixed_params), Nmif = NMIF_L) %>%
      mif2(Nmif = NMIF_L) %>%
      mif2(Nmif = NMIF_L)
    mf = mf %>%
      mif2(Nmif = NMIF_L, cooling.fraction.50 = 0.3) %>%
      mif2(Nmif = NMIF_L, cooling.fraction.50 = 0.3) %>%
      mif2(Nmif = NMIF_L, cooling.fraction.50 = 0.1) %>%
      mif2(Nmif = NMIF_L, cooling.fraction.50 = 0.1)
    ll = replicate(NREPS_EVAL, mf %>% pfilter(Np = NP) %>% logLik()) %>%
      logmeanexp(se = TRUE)
    coef(mf) %>% bind_rows() %>%
      bind_cols(loglik = ll[1],loglik.se = ll[2])
  } -> results
  attr(results,"ncpu") <- getDoParWorkers()
  results
}) %>%
  filter(is.finite(loglik)) -> results
t_global <- attr(results,"system.time")
ncpu_global <- attr(results,"ncpu")
cat(sprintf("Global search finished in %4.3f seconds\n", t_global["elapsed"]))
```


```{r, fig.height = 7, fig.width = 7}
all = read_csv(PARAMS_FILE) %>%
  filter(id <= 2) %>%
  filter(loglik > max(loglik) - 50) %>%
  bind_rows(guesses) %>%
  mutate(type = if_else(is.na(loglik), "guess", "result")) %>%
  arrange(type)

pairs(~loglik + b1 + b2 + b3 + b4 + b5, data = all,
      col = ifelse(all$type == "guess", grey(0.5), "red"), pch = 16)
pairs(~loglik + mu_EI + mu_IR + eta + rho + tau, data = all,
      col = ifelse(all$type == "guess", grey(0.5), "red"), pch = 16)
```

### Simulation results  

```{r, fig.height = 3, fig.width = 6}
optimal_params = read.csv(PARAMS_FILE) %>% slice(1) %>%
  select(-starts_with("loglik"), -id) %>% unlist()
covidSEIR %>%
  simulate(
    params = optimal_params, nsim = 10, format = "data.frame", include.data = TRUE
  ) %>%
  plot_simulation()
```
